---
title: "EarthObs_project"
author: "Ben Coombs"
date: "25/11/2020"
output: html_document
---

## Importing necessary libraries

```{r echo=FALSE}
library(openair)
library(tidyverse)
library(leaflet)
library(lubridate)
library(knitr)
library(chron)
```

## Identifying the air quality sensors in Alesund

```{r}
#Identify Norwegian air quality sensors that can be mapped from the European database
norway_meta <- importMeta(source = "Europe", all = TRUE) %>%
  filter(country == "norway", !is.na(latitude), !is.na(longitude))

#Map the sensors to identify those in Alesund
Norway_sensor_map <- leaflet(data = norway_meta) %>%
  addTiles() %>%
  addMarkers(~longitude, ~latitude, popup = ~as.character(site), label = ~as.character(site))
  
Norway_sensor_map %>%
  setView(lng = 8.468946, lat = 60.472023, zoom = 5)
```

The map above shows the locations of all ground based sensors in the European database. Unfortunately, there are no sensors available inside the World Heritage fjords. However, a popular port-of-call for cruise ships before going into the Fjords is Alesund, for which there are 3 sensors in the database. Their locations relative to the port are shown below:

```{r echo=FALSE}
Norway_sensor_map %>%
  setView(lng = 6.1495, lat = 62.4722, zoom = 14)
```

Now we need to collate the data available from all three of these sensors. The site names are only available in the metadata; in the `importEurope` dataset they are identified by the `code` variable of the metadata, so we will need to find these.

Unfortunately, we get an error when trying to get data for `Posthuskrysset`, so we'll have to leave this one out.

```{r}
Alesund_codes <- norway_meta %>%
  filter(site %in% c("Karl Eriksens plass", "Grimmerhaugen")) %>%
  pull(code)

#Get air quality data for these sites for May 2019 from European database
Alesund_air_quality <- importEurope(site = Alesund_codes,
                                    year = "2019"
                                    ) %>%
  filter( date >= as.Date("2019-05-01"), 
         date < as.Date("2019-06-01")
         )
```

For each hour, we want the `nox` measurements from site `no0116a`, and the average `pm10` measurements of the two sites. Then we can plot them to see how they behave over the month of May.

```{r}
#First select only relevant columns
Alesund_air_qual_reformatted <- Alesund_air_quality %>%
  select(-no, -no2) %>%
  pivot_wider(names_from = code, values_from = c(nox, pm10)) %>%
  select(-nox_no0070a) %>%
  mutate(
    avg_pm10 = (pm10_no0070a + pm10_no0116a)/2
  ) %>%
  rename(nox = nox_no0116a) %>%
  select(-pm10_no0070a, -pm10_no0116a) 

nox_plot <- Alesund_air_qual_reformatted%>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = nox))

pm10_plot <- Alesund_air_qual_reformatted%>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = avg_pm10))

nox_plot
pm10_plot

```

Our next mission is to create a new column `num_cruise_ships` that tells us the number of cruise ships at the port for each hour. The file `alesund_cruise_schedule.csv` contains the cruise ship schedule for 2019.

```{r}
cruises_2019_orig <- read.csv("alesund_cruise_schedule_2019.csv")

#STILL NEED TO RESOLVE NA ISSUE RATHER THAN REMOVING
cruises_2019 <- cruises_2019_orig %>%
  mutate(
    arrive = na_if(arrive, ""),
    depart = na_if(depart, "")
  ) %>%
  filter(!is.na(arrive), !is.na(depart)) %>%
  mutate(
    arrive = paste(date, arrive, sep = " "),
    depart = paste(date, depart, sep = " ")
  )

#format arrive and depart columns (couldn't get a for loop to work here either..)
cruises_2019$arrive <- as.POSIXct(cruises_2019$arrive, format = "%d-%b-%y %H:%M")
cruises_2019$depart <- as.POSIXct(cruises_2019$depart, format = "%d-%b-%y %H:%M")

cruises_2019 <- cruises_2019 %>%
  filter(arrive >= as.Date("2019-05-01 00:00:00") & depart < as.Date("2019-06-01 00:00:00"))
```

Now let's try getting the `num_cruise_ships` column

```{r}
Alesund_air_qual_ships <- Alesund_air_qual_reformatted %>%
  mutate(
    #Create col with initial value of 0
    num_cruise_ships = 0
  )

for (i in 1:nrow(Alesund_air_qual_ships)){
  #print(Alesund_air_qual_ships$date[i])
  for (j in 1:nrow(cruises_2019)){
    if (Alesund_air_qual_ships$date[i] >= cruises_2019$arrive[j] & Alesund_air_qual_ships$date[i] <= cruises_2019$depart[j]){
      Alesund_air_qual_ships$num_cruise_ships[i+1] = Alesund_air_qual_ships$num_cruise_ships[i+1] + 1
    } 
  }
}

nox_plot_ships <- Alesund_air_qual_ships %>%
  ggplot(aes(x = date, y = nox, colour = num_cruise_ships)) +
  geom_line()
nox_plot_ships

pm10_plot_ships <- Alesund_air_qual_ships %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = avg_pm10))

nox_plot_ships
pm10_plot_ships
```

